Molecular Gas Dynamics of the lensed wet-merger RXSJ1131-1123 at $z$=0.654

### 2016-05-23 WRITEUP/ms.tex
added discussion on merger

% the companion
One other interesting result from our lens model is that a better fit is
found for the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc away from the AGN host galaxy,
and C06, who find evidence for an interacting dusty galaxy.
Spatially, the red velocity component of the CO also coincides with this
component F. While we cannot quantify the type of
merger (major v.s. minor) with a mass ratio given the
spatial resolution, the symmetric velocity gradient lean support
to the case of a minor merger.

We emphasize that the spatial resolution of the data at hand
is a few arcsec, which implies that despite the high SNR and spectral
resolution, this data is insufficient to constrain the
intrinsic sizes of the lensed galaxies, and thus the magnification
factors are likely under-predicted.
We will discuss this further in the following section. % magnification factor is underpredicted, see Bussmann15

@@

% upper limit using decomposed 24 um
% lower limit using fit without 24 um, but do I still want to fit a power-law for this?
The best-fit SEDs are shown in \Fig{SED} and the resulting parameters (inferred from the BLAH of the posterior PDF) are listed in \Tab{SED}.
Our SED model find a rest-frame characteristic dust temperature of BLAH
(cf. 60$-$70\,K in the cores of local ULIRGs/SB; Pauzzo+11, Rangwala+10),
\fir (rest-frame 42.5-122.5 \micron) of blah, and a dust mass of blah.
For the mass absorption coefficient, we adopt BLAH (rest frame; Dunne et al. 2003).

@@

\subsubsection{SB versus AGN contribution}

We correct for lensing magnification using the median among our models ($\mu_{\rm L}$=BLAH).
Assuming a Salpeter initial mass function, we find the SFR of blah using the \fir luminosity after corrected for lensing .

@@

\subsubsection{Linewidth \& Sizes}
We find that the linewidths for both components (redshifted and blueshifted)
inferred from a double Gaussian fit to the line profile are
consistent with the C13 sample (FWHM of 370\,\kms). Such linewidths are higher
than in local ULIRGs (240\,\kms; \citealt[][]{Solomon97a}).
Since C13 do not have resolved observations, we compare the
gas distribution (size) with local ULIRGs. The CO gas in RXJ1131 is
distributed $\sim$6\,kpc in radius across (in
the source plane), which is more extended than the average in the
sample of disk-like ULIRGs studied by \citet{Ueda14a},
who find an average radius of 3.5\,$\pm$\,2.3\,kpc
(with a range from 1.1 kpc to 9.3 kpc).
This indicates that the positive correlation in the linewidth-size relation
holds at intermediate redshifts. Yet we cannot quantify this further
with the data represented in this paper.

\subsubsection{Gas Mass}
To derive the molecular gas mass, we first correct for lensing magnification
using the respective magnification factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing. This yields
$I$\,=\,5.9$\pm$2.7 Jy\,\kms, where the uncertainty includes those on
the magnification factors. 
We note that the derived intensity includes emission from the companion
as the spatial resolution of our data do not allow us decompose its contribution
to the total line flux.

@@

%--------------------------------------------------------------------------
%                                Discussion
%--------------------------------------------------------------------------
\section{Discussion}
% What are the conclusions of the paper (qualitative and/or quantitative)?
% What is the evidence for those conclusions?

\subsection{Mass in RXJ1131}
black hole mass
dynamical mass
stellar mass
dust mass

Assuming local $M_{\rm BH}-M_{\rm bulge}$ relation, this would imply $M_{\rm bulge}$$\sim$BLAH\Msun.

% gas-dyn fraction:
% can we quantify the merger stage of RXJ1131? what are typical gas fraction at different merger stage. What are the alpha CO used in those (0.8, see Uedaa14)
% since ellipticals are not expected to have high gas fraction. The fact that there are on-going, recent, massive episodes of SF activity, the velocity gradient, and the best-fit sersic profile from C06 all support the picture that RXJ is a disk galaxy.

\subsection{Merger stage of RXJ1131-1231}
We put our results in content with ULIRGs/mergers in $z$=0$-$1.

As discussion in previous sections, while we see evidence such as a symmetric velocity field in the source plane, double-horned profile, observed velocity
gradient, C06 model $n$=1 Sersic profile, which suggest that RXJ1131 is likely a rotating disc. Yet, this is insufficient to conclude a disk morphology. Additional characteristics such as peak velocity dispersion at the central region and a ``spider'' diagram are needed, which is commonly seen in local spirals.
\\

Kinematics and mergers:
For major mergers, two galaxy mass centers separated on the same scale as the gas disk (4 kpc or more), the gas dynamics would be strongly perturbed and would deviate substantially from that of a simple disk.

In a minor merger (a mass ratio of more than 3:1) the more massive galaxy, including its gas distribution, can retain a significant angular momentum and resemble a rotating disk.

RXJ kinematics:
The smooth, symmetric velocity field seen in our lens model and image plane
suggest a kinematically-order (i.e., disk) galaxy. However, the high
dispersion near the dynamical center suggest that RXJ1131 is turbulent, albeit
lack of spatial resolution. Overall implies RXJ1131 is a minor merger of a turbulent disc, but higher-resolution data are needed to confirm this.
Although these values are discussed in \Sec{blah} we cannot draw any general conclusions concerning BLAH.

@@

Numerical simulations of mergers between two galaxies have shown that multiple, discrete starburst events can be triggered even in double mergers. The timing of the starbursts is thought to depend on the progenitor galaxy morphologies, with early starbursts (before the galaxies start to merge) occurring in bulges disc galaxies, and late starbursts (after the galaxies have started to merge) occur- ring in disc galaxies with a bulge component (Mihos \& Hernquist 1994, 1996.
Maybe the more compact rest-frame SF and the more recent NIR SF, which is also more extended and diffuse is consistent with the spinning black hole, i.e.
already encountered?

% ------------------------------------------------------------------------
% How do these results relate to prior work?
% What are the implications of this work – i.e. what theories are being tested/questions are being answered etc?

% summarize (conflicting ) results from recent studies.
% lead way into our results

\subsection{ULIRG Evolution}
At low redshifts the ULIRG luminosity function shows strong evolution with redshift (Veilleux et al. 1999).
not known whether the IR luminosity of these high- redshift sources arises due to mergers between two or more large spiral galaxies, as in local ULIRGs, or via a different mechanism/galaxy formation processes at high-z.

gas fraction at this redshift. how does that relate to
the SFH, and the change in cosmic SFD is due to change in gas content.
SFE at this redshift, how does that relate to the SFH. probably a secondary effect to the drop in SFH

@@

% any further work needed to resolve the remaining problems

>>>>

% the companion
One other interesting result from our lens model is that a better fit is
found for the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc away from the AGN host galaxy,
and C06, who find evidence for an interacting dusty galaxy.
Spatially, the red velocity component of the CO also coincides with this
component F. While we cannot quantify the type of
merger (major v.s. minor) with a mass ratio given the
spatial resolution, the symmetric velocity gradient retained under
such close separation between the two lean support
to the idea of a minor merger.
% For major mergers, two galaxy mass centers separated on the same scale as the
% gas disk (4 kpc or more), the gas dynamics would be strongly perturbed and
% would deviate substantially from that of a simple disk.
% In a minor merger (a mass ratio of more than 3:1) the more massive galaxy, including its gas distribution, can retain a significant angular momentum and resemble a rotating disk.

We emphasize that the spatial resolution of the data in hand
is a few arcsec, which implies that despite the high SNR and spectral
resolution, this data is insufficient to constrain the
intrinsic sizes of the lensed galaxies, and thus the magnification
factors are likely under-predicted.
We will discuss this further in the following section. % magnification factor is underpredicted, see Bussmann15

@@

% upper limit using decomposed 24 um
% lower limit using fit without 24 um, but do I still want to fit a power-law for this?
The best-fit SEDs are shown in \Fig{SED} and the resulting parameters
are listed in \Tab{SED}.
Our SED model find a rest-frame characteristic dust temperature of BLAH
(cf. 60$-$70\,K in the cores of local ULIRGs/SB; Pauzzo+11, Rangwala+10),
\fir (rest-frame 42.5-122.5 \micron) of blah, and a dust mass of blah.
For the mass absorption coefficient, we adopt BLAH (rest frame; Dunne et al. 2003).

@@

\subsubsection{SB versus AGN contribution}
To derive the intrinsic properties, we correct for lensing using the 
median magnification factor derived from the CO lens models.
Assuming a Salpeter initial mass function, we find the SFR of blah using the 
lensing-corrected \fir luminosity.

@@

\subsubsection{Linewidth \& Sizes}
We find that the linewidths for both components (redshifted and blueshifted)
inferred from a double Gaussian fit to the line profile are
consistent with the C13 sample (FWHM of 370\,\kms). Such linewidths are higher
than local ULIRGs (240\,\kms; \citealt[][]{Solomon97a}).
Since C13 do not have resolved observations, we compare the
gas distribution with local ULIRGs. The CO gas in RXJ1131 is
distributed $\sim$6\,kpc in radius across (in
the source plane), which is more extended than the average in the
sample of disk-like ULIRGs studied by \citet{Ueda14a},
who find an average radius of 3.5\,$\pm$\,2.3\,kpc
(with a range from 1.1 kpc to 9.3 kpc).
This indicates that the positive correlation in the linewidth-size relation
holds at intermediate redshifts. Yet we cannot quantify this further
with the data represented in this paper.

\subsubsection{Gas Mass}
To derive the molecular gas mass, we first correct for lensing magnification
using the respective magnification factors listed in \Tab{model}, which to
first order takes into account effect of differential lensing. This yields
$I_{\tiny \bco}$\,=\,5.9$\pm$2.7 Jy\,\kms, where the uncertainty includes those on
the magnification factors.
We note that the derived intensity includes emission from the companion
as the spatial resolution of our data do not allow us decompose its contribution
from the total line flux.

@@

%--------------------------------------------------------------------------
%                                Discussion
%--------------------------------------------------------------------------
\section{Discussion}
\subsection{Merger stage of RXJ1131-1231}
In this section, we put our results in context with other ULIRGs/mergers at
$z$\,=\,0$-$1. It is clear that RXJ1131 has a gas-rich companion, but how does
this affect its internal dynamics and what is the merger stage of RXJ1131?
We argued that the symmetric source-plane velocity field suggest
the system is a minor merger, consistent with an earlier study by B08, who
independently conclude that the companion is likely
a dwarf of size $\sim$700\,kpc across.
While our result is consistent with previous studies of RXJ1131, this
contrast strongly with studies of mergers, where powerful AGN are
usually found in late stage mergers (Yuan+10, Iwasawa+11) and
that only major mergers near their final
coalescence can power/provide \LIR\,$>$\,10$^{12}$\,\Lsun, based on
a recent study of a sample of 65 local U/LIRGs
(\LIR\,=\,10$^{11-12.5}$\,\Lsun; $z$\,$<$\,0.1; Larson+16).
If this is also the case at $z$$\sim$0.65, this would imply RXJ1131
is a major merger near its final coalescence and that the spatially offset
component is an expulsion of gas driven by AGN winds or is some
tidally ejected material from previous passage.
In that case, the molecular gas fraction in the host
galaxy should have declined (as in ellipticals) compared to early
stage mergers (can we find molecular gas mass fraction as a
function of merger stage?) and that the molecular gas would
be more concentrated in the nuclear region.
This is inconsistent with what
is observed (which strongly favors a disk morphology; 
but there are outliers --  Ueda+14 find
extended molecular gas in merger remnant) and
the fact that there is on-going \SF activity.
Hence, we reject this and speculate the discrepancy is a consequence
of (the power source of) IR luminous galaxies are already very different by this redshift.
An alternative explanation might be that the
CO is less strongly-magnified as compared to the dust emission,
in which case, the lensing-corrected \LFIR is overestimated.
This is plausible given the extent of the source with respect to the caustics
and that substantial differential lensing effect has been seen across CO and other
wavelengths. If we instead use the NIR ($H$-band) magnification factor ($\mu_{\rm L}$\,=\,BLAH; C06), 
we find \LFIR\,=\,BLAH. This would imply that RXJ1131 is consistent with a minor
merger in the context of the results reported by Larson+16.

\subsubsection{Fate of RXJ1131-1231}
Has RXJ1131 encountered with the companion?

@@

Numerical simulations of mergers have shown that
multiple starburst episodes can be triggered even in double mergers.
The timing of the starbursts is thought to depend on the progenitor galaxy
morphologies, with early starbursts (before the galaxies start to merge)
occurring in bulges disc galaxies, and late starbursts (after the galaxies
have started to merge) occurring in disc galaxies with a bulge component (Mihos \& Hernquist 1994, 1996.
The spinning black hole already encountered?

The future of RXJ1131:
classical picture is that merger will evolve into local ellipticals, but more
recent studies (springel \& Hernquist 05, cite) suggest that mergers can also
form local large-disks (bulge dominated) $<--$ double check this. Given the
extended molecular gas distribution, we suggest RXJ1131 may evolve into the
latter (large-disk) unless the ``companion'' BKAH SOME STUFF or some other
mechanisms at play in the latter stage, transporting the gas
toward the central nuclear region.

\subsection{ULIRG Evolution and Cosmic SFH}
At low redshifts, the luminosity function of ULIRG shows strong evolution with
redshift (Veilleux et al. 1999). It is not known whether the IR luminosity of
these high-$z$ sources arises due to mergers between two or more large spiral
galaxies, as in local ULIRGs, or via a different mechanism/galaxy formation
processes at high-z.

% AGN contribution.

% Implication on SFH
Our result indicates that the gas mass fraction has already decreased
at this redshift, which may suggest that the change in cosmic SFD is due to
change in gas content. The SFE is also BLAH at this redshift,
consistent with current speculation of how the SFRD decreases over cosmic time.

@@

% any further work needed to resolve the remaining problems
We list several pieces of evidence through out
this work to argue that RXJ1131 is likely a rotating disk.
The double-horned profile, smooth, symmetric velocity field
seen in our lens model and in the
image plane and C06 model $n$=1 Sersic profile all suggest
a kinematically-order (i.e., disk) galaxy.
Yet, these are insufficient to conclude a disk morphology. Additional
characteristics such as peak velocity dispersion at the central region and a
``spider'' diagram are needed, which is commonly seen in local spirals.
The high dispersion near the dynamical center also
suggest that RXJ1131 is turbulent, albeit lack of spatial resolution. Overall
implies RXJ1131 is a minor merger of a turbulent disc, but higher-resolution
data are needed to confirm this.
